---
title: "Progress Memo 2"
subtitle: |
  | Final Project 
  | Data Science 2 with R (STAT 301-2)
author: "Karthik Sathish"
date: today

format:
  html:
    toc: true
    embed-resources: true
    
execute:
  echo: false
  warning: false

from: markdown+emoji 
reference-location: margin
citation-location: margin
---



::: {.callout-tip icon=false}

## Github Repo Link

[My Github Repo Link](https://github.com/stat301-2-2024-winter/final-project-2-kvsathish.git)

:::

::: {.callout-warning}
GitHub (free account) cannot store large files. If you try to commit and push a large dataset you will have an ERROR! Any file over 100 MB (100,000 KB) needs to be added to the `.gitignore` file BEFORE committing.

**We may need to do that for files in your final project!**
:::

```{r}
#| echo: false

# load packages
library(tidyverse)
library(skimr)
library(ggplot2)
library(here)
library(kableExtra)
library(knitr)
library(naniar)

load("memo_data/bball_players.rda")
```



## New Analysis of Target Variable

```{r}
#| echo: false

bball_players |> 
  ggplot(aes(x = pick)) +
  geom_bar() +
  labs(title = "Distribution of `pick`",
       y = NULL) +
  theme_minimal()
```


## Assessment Metric

????

do u change character to factor vars?

ask about gitignore?

Blah blah blah^[an example of footnote --- [content in margins](https://quarto.org/docs/blog/posts/2022-02-17-advanced-layout/)]

## Analysis Plan

Blah blah blah^[another footnote]

## Modelling

Blah blah blah

## Metric Evaluation of Models

Blah blah blah

## Progress

Blah Blah
